DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-08-30 2015CompleteDerivation/completeDerivation.tex
Using namespace for labels?

\begin{assump}[Determinism and reversibility]\label{detrevass}
The state of the physical system under study undergoes deterministic and reversible evolution.
\end{assump}

@@

\begin{prop}\label{differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a  classical material is a differentiable manifold. The state space $\mathcal{C}$ for a classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$. Let $q^i$ and $\hat{q}^j$ two sets of independent state variables in $U\subseteq \mathcal{S}$. Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.  $\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as $\rho_\mathcal{c}$ transforms as a density. $\rho_\mathcal{c}(\mathcal{s}) = \rho_\mathcal{c}(q^i(\mathcal{s})) = \rho_\mathcal{c}(\hat{q}^j(\mathcal{s}))$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

@@

\begin{assump}[Infinitesimal reducibility]\label{infinitesimal_reducibility}
	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
\end{assump}

@@

	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist. Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption. The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with molecules, atoms and then subatomic particles. But it is more instructive to understand when and how the assumption breaks down at a more conceptual level.

@@

	The second, more conceptual, problem is that the idealizations that allowed us to define states in Assumption \ref{detrevass} may not hold as the parts get smaller. Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface. As we consider smaller and smaller parts, at some point we will find an amount of material that is affected by the interaction with the air or photons scattering off of it. At that point the parts are no longer sufficiently isolated to define an independent state, their evolution depends on the environment. Therefore the deterministic evolution of the whole cannot be reduced to the deterministic evolution of its parts.

@@

\begin{justification}
	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$. Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$. Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable. 
	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{differentiable_manifold}. $\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}. $\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables. $\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$

	We claim $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$. As in \ref{differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
\end{justification}

@@

	We claim $\bar{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$. $\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities. Consider the map $f : T^*\mathcal{M} \rightarrow \bar{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$. $f$ is an open map. The topology of $\bar{\mathcal{S}}$ is the subspace topology. Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density. $\rho$ is differentiable by \ref{differentiable_manifold}. It's restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable. The inclusion map $\bar{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding. $\bar{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$. $\bar{\mathcal{S}}$ is an hypersurface.
\end{justification}

@@

\begin{prop}\label{antistates}
	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time dependent deterministic and reversible evolution map for the particles of a classical material. The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
\end{prop}

@@

\begin{assump}[Kinematic equivalence]\label{kinematic_equivalence}
	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
\end{assump}

@@

	We claim $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism. Let $\bar{\mathcal{c}}$ be the evolution of a minimal classical material. Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution associated with $\bar{\mathcal{c}}$. $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal. $\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{differentiable_manifold}. $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
\end{justification}

@@

The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible. These correspond to the particle anti-states mentioned in \ref{antistates}. As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$. The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard particle state or an anti-state. Expanding the equation for the force, we have:
\begin{align*}
\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
\end{align*}
Therefore the force acts on anti-particle states in the opposite way: they act as if they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.

@@

Assuming such a background is not far fetched. In fact, it serves as a remainder of the issues mentioned in assumption \ref{infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated. If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each particle is not negligible.

>>>>

\begin{assump}[Determinism and reversibility]\label{ass:determinism}
The state of the physical system under study undergoes deterministic and reversible evolution.
\end{assump}

@@

\begin{prop}\label{prop:differentiable_manifold}
	The state space $\mathcal{S}$ for the particles of a  classical material is a differentiable manifold. The state space $\mathcal{C}$ for a classical material is isomorphic to the space of Lebesgue integrable differentiable functions. That is $\mathcal{C} \cong C^1(\mathcal{S}) \cap L^1(\mathcal{S}, \mu)$, where $C^1(\mathcal{S}) \equiv \{\rho:\mathcal{S} \rightarrow \mathbbm{R} \; | \; \rho$ is differentiable$\}$.
\end{prop}
\begin{justification}
	We claim $\mathcal{S}$ is a differential manifold. Let $\rho_\mathcal{c} : \mathcal{S} \rightarrow \mathbbm{R}$ be the distribution associated with a composite state $\mathcal{c} \in \mathcal{C}$. Let $q^i$ and $\hat{q}^j$ two sets of independent state variables in $U\subseteq \mathcal{S}$. Let $\rho_\mathcal{c}(q^i)$ and $\rho_\mathcal{c}(\hat{q}^j)$ be the expression in local coordinates of the distribution.  $\rho_\mathcal{c}(\hat{q}^j)=\rho_\mathcal{c}(q^i) | \partial_j q^i |$ as $\rho_\mathcal{c}$ transforms as a density. $\rho_\mathcal{c}(\mathcal{s}) = \rho_\mathcal{c}(q^i(\mathcal{s})) = \rho_\mathcal{c}(\hat{q}^j(\mathcal{s}))$. The Jacobian $| d q' / d q |$ exists and is non-zero. The map between any two charts is differentiable. $\mathcal{S}$ is a differentiable manifold.

@@

\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
	The system under study is composed of an infinitesimally reducible homogeneous material and each part undergoes deterministic and reversible evolution.
\end{assump}

@@

	Consistently with what we said in assumption \ref{ass:determinism}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist. Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption. The obvious reason is that we know this does not work in practice: as we keep decomposing the material we end up with molecules, atoms and then subatomic particles. But it is more instructive to understand when and how the assumption breaks down at a more conceptual level.

@@

	The second, more conceptual, problem is that the idealizations that allowed us to define states in assumption \ref{ass:determinism} may not hold as the parts get smaller. Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface. As we consider smaller and smaller parts, at some point we will find an amount of material that is affected by the interaction with the air or photons scattering off of it. At that point the parts are no longer sufficiently isolated to define an independent state, their evolution depends on the environment. Therefore the deterministic evolution of the whole cannot be reduced to the deterministic evolution of its parts.

@@

\begin{justification}
	We claim $\bar{\mathcal{C}} \cong  G \subseteq C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$. Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution of material associated with an evolution $\bar{\mathcal{c}} \in \bar{\mathcal{C}}$. Let $t : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be a time variable. 
	The restriction of $\rho_{\bar{\mathcal{c}}}$ over an arbitrary level set $\mathcal{S}_{t=t_0}$ is continuous and integrable by \ref{prop:differentiable_manifold}. $\rho_{\bar{\mathcal{c}}} \in L^1_t(\bar{\mathcal{S}}, \mu)$ Time evolution is a diffeomorphism by \ref{hamiltons_equations}. $\rho_{\bar{\mathcal{c}}}$ is differentiable along the time variable as well as the state variables. $\rho_{\bar{\mathcal{c}}} \in C^1(\bar{\mathcal{S}})$

	We claim $\bar{\mathcal{C}} \cong  C^1(\bar{\mathcal{S}})\cap L^1_t(\bar{\mathcal{S}}, \mu)$. As in \ref{prop:differentiable_manifold}, if two distributions are different there exists an outcome that can tell them apart, therefore they represent two distinct physical evolutions.
\end{justification}

@@

	We claim $\bar{\mathcal{S}}$ is a hyper-surface of $(T^*\mathcal{M}, \omega)$. $\bar{\mathcal{S}} \subseteq (T^*\mathcal{M}, \omega)$ as states are physical objects that allow time-state-invariant-variable densities. Consider the map $f : T^*\mathcal{M} \rightarrow \bar{\mathcal{S}} \; | \; (q^i, p_i, t, E) \mapsto (q^i, p_i, t)$. $f$ is an open map. The topology of $\bar{\mathcal{S}}$ is the subspace topology. Let $\rho : T^*\mathcal{M} \rightarrow \mathbbm{R}$ be an invariant density. $\rho$ is differentiable by \ref{prop:differentiable_manifold}. It's restriction on $\bar{\mathcal{S}}$ is also an invariant distribution and therefore differentiable. The inclusion map $\bar{\mathcal{S}} \hookrightarrow T^*\mathcal{M}$ is a smooth embedding. $\bar{\mathcal{S}}$ is an embedded submanifold of co-dimension $1$. $\bar{\mathcal{S}}$ is an hypersurface.
\end{justification}

@@

\begin{prop}\label{prop:antistates}
	Let $\mathcal{T}_{\Delta s}: T^*\mathcal{M} \rightarrow T^*\mathcal{M}$ the time dependent deterministic and reversible evolution map for the particles of a classical material. The map partitions the extended state space into standard states, those connected by a trajectory where $d_{s}t>0$, and anti-states, those connected by a trajectory where $d_{s}t<0$.
\end{prop}

@@

\begin{assump}[Kinematic equivalence]\label{ass:kinematic_equivalence}
	For the given classical material, studying its kinematics  (i.e. trajectories in space) is equivalent to study its dynamics  (i.e. states and their evolution).
\end{assump}

@@

	We claim $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism. Let $\bar{\mathcal{c}}$ be the evolution of a minimal classical material. Let $\rho_{\bar{\mathcal{c}}} : \bar{\mathcal{S}} \rightarrow \mathbbm{R}$ be the distribution associated with $\bar{\mathcal{c}}$. $\rho_{\bar{\mathcal{c}}}(\mathcal{s}(x^\alpha, u^\alpha))$ as the material is minimal. $\rho_{\bar{\mathcal{c}}}(x^\alpha, u^\alpha)$ is differential for the same arguments expressed in \ref{prop:differentiable_manifold}. $(x^\alpha, u^\alpha) \mapsto (q^\alpha, p_\alpha)$ is a diffeomorphism as it as a homeomorphism that maps a differentiable function to a differentiable function.
\end{justification}

@@

The extended Hamiltonian we found is quadratic in energy, therefore negative energy states are possible. These correspond to the particle anti-states mentioned in \ref{prop:antistates}. As we fixed the natural parametrization, $u^\alpha u_\alpha = - c^2$, there are two cases: the parametrization is aligned with time, therefore $s=\tau$ where $\tau$ is proper time, or the parametrization is anti-aligned with time, therefore $s=-\tau$. The trajectory velocity $u^\alpha= \pm d_\tau x^\alpha$ is plus or minus the physical velocity depending on whether it's a standard particle state or an anti-state. Expanding the equation for the force, we have:
\begin{align*}
\pm d_\tau (\pm m d_\tau x^\alpha) &= m d_\tau^2 x^\alpha = \pm  F^\alpha_\beta d_\tau x^\beta
\end{align*}
Therefore the force acts on anti-particle states in the opposite way: they act as if they have opposite charge. Note that nothing physical is actually going backwards in time, just the parametrization.

@@

Assuming such a background is not far fetched. In fact, it serves as a remainder of the issues mentioned in assumption \ref{ass:infinitesimal_reducibility}: as determinism and reversibility at all scales conceptually does not work, there exists a minimum constant interaction to all particles: the system is not perfectly isolated. If the entropy of our distribution is comparable to the entropy introduced by the background, we are likely in the range where classical mechanics is not valid as the random motion of each particle is not negligible.
